---
title: "AnnData to Seurat"
author: "Bence Kover"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the required libraries

```{r}
library(anndataR)
library(Seurat)
library(ggplot2)
options(warn = -1)
```
# RNA

Load in the AnnData objects we have annotated in Python using epitome_tools

```{r}
adata <- read_h5ad("mouse_multiome_rna.h5ad")

seurat_rna <- adata$as_Seurat(
  assay_name = "RNA",
  layers_mapping = c(counts = "raw_counts")
)
```

Setting a cell type palette for plotting to match the plots we generated in Python.

```{r}
cell_type_palette <- c(
  Corticotrophs      = "#1f77b4",
  Endothelial_cells  = "#ff7f0e",
  Erythrocytes       = "#2ca02c",
  Gonadotrophs       = "#d62728",
  Immune_cells       = "#9467bd",
  Lactotrophs        = "#8c564b",
  Melanotrophs       = "#e377c2",
  Mesenchymal_cells  = "#7f7f7f",
  Pituicytes         = "#bcbd22",
  Somatotrophs       = "#17becf",
  Stem_cells         = "#aec7e8",
  Thyrotrophs        = "#ffbb78"
)
```

We can now set the identity (cluster naming) of cells to cell_type_final from the epitome_tools output
We then proceed to plot clusters and stem cell probability scores on a UMAP

```{r}
Idents(seurat_rna) <- seurat_rna@meta.data$cell_type_final

#Plotting clusters
p1 <- DimPlot(seurat_rna, cols=cell_type_palette )
#Plotting stem cell probability
p2 <- FeaturePlot(
  seurat_rna,
  features = "proba_Stem_cells"
)

# Save as PNG
ggsave("DimPlot_clusters.png", plot = p1,
       width = 6.5, height = 4, dpi = 800)

ggsave("FeaturePlot_proba_Stem_cells.png", plot = p2,
       width = 6.5, height = 4, dpi = 800)
```

#ATAC

Loading in the ATAC data requires starting again from the fragment file
For peaks we will use the Consensus Chromatin Landscape once more

```{r}
library(Signac)
library(Seurat)
library(EnsDb.Mmusculus.v79)
library(GenomicRanges)
library(GenomeInfoDb)
library(Rsamtools)

frag_path <- "/Users/bencekover/Downloads/fragments.tsv.gz"
indexTabix(frag_path, format = "bed")
#https://raw.githubusercontent.com/Andoniadou-Lab/consensus_pituitary_atlas/main/Consensus_Chromatin_Landscape/consensus_chromatin_landscape.bed
peak_path <- "/Users/bencekover/Downloads/consensus_chromatin_landscape.bed"
```


```{r}
# annotation (mm10, UCSC-style chr names)
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"

# read consensus peaks
peaks <- read.table(peak_path, col.names = c("chr", "start", "end"))
peaks <- makeGRangesFromDataFrame(peaks)

# fragment object + quantify consensus peaks x cells
frags  <- CreateFragmentObject(path = frag_path)
counts <- FeatureMatrix(fragments = frags, features = peaks)
```

```{r}
library(anndataR)
library(Seurat)

adata <- read_h5ad("mouse_multiome_atac_final.h5ad")

python_atac <- adata$as_Seurat(
  assay_name = "ATAC",
)

#keep only cells that occur in python_atac
common_cells <- intersect(colnames(counts), colnames(python_atac))
```
We can now generate a new Signac object, and transfer over the metadata

```{r}
counts <- counts[, common_cells]

# build Signac object
signac_atac <- CreateSeuratObject(
  counts = CreateChromatinAssay(
    counts     = counts,
    fragments  = frags,
    annotation = annotation,
    genome     = "mm10"
  ),
  assay = "ATAC"
)

#transferring over the metadata
signac_atac@meta.data <- python_atac@meta.data

#transferring over UMAP coords
signac_atac@reductions$umap <- python_atac@reductions$X_umap


```

We can now set idents (cluster names) and generate plots.
```{r}
Idents(signac_atac) <- signac_atac@meta.data$cell_type_final

#Plotting clusters
p1 <- DimPlot(signac_atac, cols=cell_type_palette )
#Plotting stem cell probability
p2 <- FeaturePlot(
  signac_atac,
  features = "proba_Stem_cells"
)

# Save as PNG
ggsave("DimPlot_clusters_atac.png", plot = p1,
       width = 6.5, height = 4, dpi = 800)

ggsave("FeaturePlot_proba_Stem_cells_atac.png", plot = p2,
       width = 6.5, height = 4, dpi = 800)


```



